Summary of the things that I learned

### 2024-11-25 200-category-theory.tex
Sort out simeq and cong

@@

Alternatively, we can characterize universal arrows by their action on hom-sets:
\begin{lemma}
  Let $ F: \C \to \D $ be a functor and $ X: \D $ an object. An object $ Y: \C $ and an arrow $ f: \D(X, F(Y)) $ form a universal arrow from $ X $ to $ F $ if and only if the function
  \[ (g \mapsto f \cdot F(g)) : \C(Y, x) \cong \D(X, F(x)) \]
  is a bijection.

@@

Now, this embedding has a couple of properties:
\begin{lemma}
  For any $ Y: \C $ and $ F : P \C $, we have an equivalence $ P \C(\yo(Y), F) \simeq F(Y) $, and this equivalence is natural in $ Y $ and $ F $.
\end{lemma}
\begin{proof}
  It sends a natural transformation $ a: P \C(\yo(Y), F) $ to the element $ a_Y(\id Y) $. Conversely, it sends an element $ X : P(Y) $ to the natural transformation $ a $ given by $ a_X(f) = P(f)(X) $ for all $ f: \yo(Y)(X) = \C(X, Y) $. For more details, see \autocite[][Proposition 1.4.3]{Kashiwara}.
\end{proof}

@@

In fact, $ \Delta_X : \C \to (\C \downarrow X) $ is a functor, with $ \Delta_X(f) = \id X \times f $ for $ f: \C(Y, Y^\prime) $. This functor preserves the terminal object, products and pullbacks:
\begin{lemma}\label{lem:delta-limits}
  $ \Delta_X $ preserves all limits.
\end{lemma}
\begin{proof}
  Take a diagram $ ((Y_i)_i, (f_j)_j) $ in $ \C $. Suppose that this has a limit $ L : \C $ with projections $ g_i : \C(L, Y_i) $. Now, consider an object $ (Z, q) : (\C \downarrow X) $, together with morphisms $ h_i : (\C \downarrow X)((Z, q), \Delta_X Y_i) $, that commute with the $ \Delta_X f_j $:
  \begin{center}
    \begin{tikzcd}
      & Z \arrow[ldd, "h_{m_j}"'] \arrow[rdd, "h_{n_j}"]\\
      & X \times L \arrow[ld, "\Delta_X g_{m_j}"] \arrow[rd, "\Delta_X g_{n_j}"']\\
      X \times Y_{m_j} \arrow[rr, "\Delta_X f_j"'] & & X \times Y_{n_j}
    \end{tikzcd}
  \end{center}
  Then the morphisms in $ (\C \downarrow X)((Z, q), \Delta_X L) $, commuting with the $ \Delta_X g_j $ and $ h_j $ are the morphisms in $ \C(Z, X \times L) \cong \C(Z, X) \times \C(Z, L) $ that commute with the projections to $ X $ and the $ \id X \times g_j $ and $ h_j $. Since the morphisms in $ \C(Z, X) $ commuting with $ q $ and $ \id X $ are exactly $ q $, we can forget about this component, and the morphisms we are looking for correspond to the morphisms in $ \C(Z, L) $ that commute with the $ g_j $ and $ h_j \cdot p_1 $. Since $ (L, (g_j)_j) $ is a limit, this is a unique morphism.
\end{proof}

\begin{lemma}\label{lem:delta-exponentials}
  $ \Delta_X $ preserves exponential objects:
\end{lemma}
\begin{proof}
  See \autocite[][Volume 3, Lemma 5.8.2]{borceux}. There, the following equivalences are used for the proof:
  \begin{align*}
    (\C \downarrow X)((W, f), \Delta_X Y^Z) &\cong \C(W, Y^Z)\\
    &\cong \C(W \times Z, Y)\\
    &\cong (\C \downarrow X)((W \times (X \times Z), \langle f, p_1 \rangle), \Delta_X Y)\\
    &\cong (\C \downarrow X)((W, f) \times \Delta_X Z, \Delta_X Y)\\
  \end{align*}
\end{proof}

@@

Now, in $ \SET $, for a family $ (Y_x)_x $, consider the dependent product $ \prod_{x : X} Y_x $. Its elements $ (y_x)_x $ can be identified with morphisms from the terminal set: $ \{ \star \} \to \prod_{x : X} Y_x $, sending $ \star $ to $ (y_x)_x $. However, they can also be identified with the morphisms $ f: X \to \sum_{x : X} Y_x $ that make the following diagram commute, sending $ x $ to $ y_x $:
\begin{center}
  \begin{tikzcd}
    X \arrow[rd, "\id X"'] \arrow[rr, "f"] & & \sum_{x : X} Y_x \arrow[ld, "\pi_1"]\\
    & X
  \end{tikzcd}
\end{center}
These are morphisms in $ (\SET \downarrow X) $ from $ (X, \id X) $ to $ (\sum_{x : X} Y_x, \pi_1) $. Note that for the terminal morphism $ f: X \to \{ \star \} $, we have $ (X, \id X) = f^*(\{\star\}, \id{\{\star\}}) $. To summarize, we have an equivalence
\[ (\SET \downarrow X)\left(f^*(\{ \star \}, \id{\{\star\}}), (Y_x)_x \right) \simeq (\SET \downarrow {\{\star\}})\left( (\{\star\}, \id{\{ \star \}}), \left( \prod_{x : X} Y_x, ! \right) \right) \]
for $ ! $ the terminal morphism. Now, for an internal indexed family $ f : Y \to X $ and given a family of families $ ((Z_y)_{y : Y_x})_{x : X} $, we can wonder whether we can construct the family of dependent products $ (\prod_{y : Y_x} Z_y)_x $. In $ \SET $, this is definitely possible, and from this, we get an equivalence again
\[ \left(\SET \downarrow \sum_{x : X} Y_x \right)\left(f^*(X, \id X), (Z_y)_{x, y} \right) \simeq (\SET \downarrow X)\left((X, \id X), \left(\prod_{y : Y_x} Z_y\right)_x\right). \]
with $ p : \sum_{x : X} \sum_{y : Y_x} Z_y \to \sum_{x : X} Y_x $ defined as $ p(x, y, z) = (x, y) $. These equivalences suggest an adjunction $ f^* \dashv \prod_{\dots} $. We can use this to define in general
\begin{definition}
  For a category $ \C $ and a morphism $ f: Y \to X $, the \iindex{dependent product} along $ f $ is, if it exists, the right adjoint to the pullback functor:
  \begin{center}
    \begin{tikzcd}
      (\C \downarrow X) \arrow[r, shift left=2, "f^*"{name=A}] &
      (\C \downarrow Y) \arrow[l, shift left=2, "\prod_f"{name=B}]
      \ar[from=A, to=B, symbol=\dashv]
    \end{tikzcd}
  \end{center}
\end{definition}
\begin{remark}
  As argued above, we can recover the familiar dependent product $ \prod_{x : X} Y_x $ of a family $ (Y_x)_x $ as the dependent product $ \prod_f (\sum_x Y_x, \pi_1) $ along the terminal morphism $ f: X \to I $. Here we use the equivalence between $ (\C \downarrow I) $ and $ \C $.
\end{remark}

Now we turn our attention to dependent sums. In $ \SET $, let $ (Y_x)_x $ and $ (Y^\prime_x)_x $ be two families over $ X $ and let $ ((Z_y)_{y : Y_x})_{x : X} $ be a family of families. Let $ f: \sum_{x : X} Y_x \to X $ be the internal representation of $ (Y_x)_x $. A family of maps $ g_x : (\sum_{y : Y_x} Z_y)_x \to Y^\prime_x $ consists of maps $ Z_y \to Y^\prime_x $ for all $ y : Y_x $, so these are maps $ g_y : Z_y \to Y^\prime_{f(y)} $. This gives an equivalence
\[ (\SET \downarrow X)\left(\left( \sum_{y : Y_x} Z_y \right)_x, (Y^\prime_x)_x \right) \simeq \left(\SET \downarrow \left(\sum_{x : X} Y_x \right) \right)((Z_y)_y, f^*((Y^\prime_x)_x)). \]
This, again, suggests an adjunction which we will use as a definition.
\begin{definition}
  For a category $ \C $ and a morphism $ f: Y \to X $, the \iindex{dependent sum} along $ f $ is, if it exists, the left adjoint to the pullback functor:
  \begin{center}
    \begin{tikzcd}
      (\C \downarrow X) \arrow[r, shift right=2, "f^*"'{name=B}] &
      (\C \downarrow Y) \arrow[l, shift right=2, "\sum_f"'{name=A}]
      \ar[from=A, to=B, symbol=\dashv]
    \end{tikzcd}
  \end{center}
\end{definition}

However, note that the conversion from an external to an internal representation in $ \SET $ already contained a dependent sum, which is no coincidence. It turns out that in practice, we will never have a hard time obtaining dependent sums:
\begin{lemma}\label{lem:sum-postcomposition}
  Let $ f : \C(Y, X) $ be a morphism in a category. If the pullback functor $ f^*: (\C \downarrow X) \to (\C \downarrow Y) $ exists, it has a left adjoint given by postcomposition with $ f $.
\end{lemma}
\begin{proof}
  For morphisms $ g: Z \to X $, $ h: W \to Y $, the universal property of the pullback, with the following diagram
  \begin{center}
    \begin{tikzcd}
      W \arrow[r, dashed, "\varphi"'] \arrow[rr, "\psi", bend left] \arrow[dr, "h"'] & f^* Z \arrow[d, "f^* g"'] \arrow[r] \arrow[rd, phantom, "\lrcorner", very near start] & Z \arrow[d, "g"]\\
      & Y \arrow[r, "f"'] & X
    \end{tikzcd}
  \end{center}
  gives an equivalence between morphisms $ \varphi: W \to f^* Z $ that commute with $ h $ and $ f^* g $, and morphisms $ \psi: W \to Z $ that commute with $ h $, $ g $ and $ f $. In other words:
  \[ (\C \downarrow X)(h \cdot f, g) \simeq (\C \downarrow Y)(h, f^*(g)), \]
  which shows the adjunction.
\end{proof}

@@

  Now, note that if we have a coequalizer $ W $ of $ \id Z $ and $ a $, and an equalizer $ Y $ of $ \id X $ and $ b $ (in particular, if $ W $ and $ Y $ are retracts), the universal properties of these give an equivalence
  \[ \D(W, Y) \cong \{ f: \D(Z, Y) \mid a \cdot f = f \} \cong \{ f: \D(Z, X) \mid a \cdot f = f = f \cdot b \}. \]
  \begin{center}
    \begin{tikzcd}
      Z \arrow[r, shift left, "\id Z"] \arrow[r, shift right, "a"'] & Z \arrow[r] \arrow[d] \arrow[rd] & W \arrow[d]\\
      X & X \arrow[l, shift left, "\id X"] \arrow[l, shift right, "b"'] & Y \arrow[l]
    \end{tikzcd}
  \end{center}
\end{proof}

@@

\begin{corollary}\label{cor:karoubi-presheaf}
  As the category $ \SET $ is cocomplete, we have an equivalence between the category of presheaves on $ \C $ and the category of presheaves on $ \overline \C $:
  \[ [\op \C, \SET] \cong [\overline{\op \C}, \SET] \cong [\op{\overline \C}, \SET]. \]
\end{corollary}

>>>>

Throughout this thesis, we will use the notation $ A \cong B $ for an isomorphism of objects in a category, or of categories themselves. We will use the notation $ A \simeq B $ for an equivalence of types or of categories. If $ A $ and $ B $ are sets or univalent categories, $ A \cong B $ is equivalent to $ A \simeq B $. We will often, instead of $ A \cong B $ or $ A \simeq B $ just write $ f: A \xrightarrow \sim B $, giving the isomorphism or equivalence a name.

@@

Alternatively, we can characterize universal arrows by their action on hom-sets:
\begin{lemma}
  Let $ F: \C \to \D $ be a functor and $ X: \D $ an object. An object $ Y: \C $ and an arrow $ f: \D(X, F(Y)) $ form a universal arrow from $ X $ to $ F $ if and only if the function
  \[ (g \mapsto f \cdot F(g)) : \C(Y, x) \to \D(X, F(x)) \]
  is a bijection.

@@

Now, this embedding has a couple of properties:
\begin{lemma}
  For any $ Y: \C $ and $ F : P \C $, we have a bijection $ P \C(\yo(Y), F) \cong F(Y) $, and this bijection is natural in $ Y $ and $ F $.
\end{lemma}
\begin{proof}
  It sends a natural transformation $ a: P \C(\yo(Y), F) $ to the element $ a_Y(\id Y) $. Conversely, it sends an element $ X : P(Y) $ to the natural transformation $ a $ given by $ a_X(f) = P(f)(X) $ for all $ f: \yo(Y)(X) = \C(X, Y) $. For more details, see \autocite[][Proposition 1.4.3]{Kashiwara}.
\end{proof}

@@

In fact, $ \Delta_X : \C \to (\C \downarrow X) $ is a functor, with $ \Delta_X(f) = \id X \times f $ for $ f: \C(Y, Y^\prime) $. If $ \C $ has a terminal object $ T $ and we identify $ \C $ with $ (\C \downarrow T) $, then $ \Delta_X $ is the pullback functor along $ (\C \downarrow T) \to (\C \downarrow X) $ along the morphism $ !: X \to T $. This functor preserves the terminal object, products and pullbacks:
\begin{lemma}\label{lem:delta-limits}
  $ \Delta_X $ preserves all limits.
\end{lemma}
\begin{proof}
  Take a diagram $ ((Y_i)_i, (f_j)_j) $ in $ \C $. Suppose that this has a limit $ L : \C $ with projections $ g_i : \C(L, Y_i) $. Now, consider an object $ (Z, q) : (\C \downarrow X) $, together with morphisms $ h_i : (\C \downarrow X)((Z, q), \Delta_X Y_i) $, that commute with the $ \Delta_X f_j $:
  \begin{center}
    \begin{tikzcd}
      & Z \arrow[ldd, "h_{m_j}"'] \arrow[rdd, "h_{n_j}"]\\
      & X \times L \arrow[ld, "\Delta_X g_{m_j}"] \arrow[rd, "\Delta_X g_{n_j}"']\\
      X \times Y_{m_j} \arrow[rr, "\Delta_X f_j"'] & & X \times Y_{n_j}
    \end{tikzcd}
  \end{center}
  Then the morphisms in $ (\C \downarrow X)((Z, q), \Delta_X L) $, commuting with the $ \Delta_X g_j $ and $ h_j $ are the morphisms in $ \C(Z, X \times L) \cong \C(Z, X) \times \C(Z, L) $ that commute with the projections to $ X $ and the $ \id X \times g_j $ and $ h_j $. Since the morphisms in $ \C(Z, X) $ commuting with $ q $ and $ \id X $ are exactly $ q $, we can forget about this component, and the morphisms we are looking for correspond to the morphisms in $ \C(Z, L) $ that commute with the $ g_j $ and $ h_j \cdot p_1 $. Since $ (L, (g_j)_j) $ is a limit, this is a unique morphism.
\end{proof}

\begin{lemma}\label{lem:delta-exponentials}
  $ \Delta_X $ preserves exponential objects:
\end{lemma}
\begin{proof}
  See \autocite[][Volume 3, Lemma 5.8.2]{borceux}. There, the following isomorphisms are used for the proof:
  \begin{align*}
    (\C \downarrow X)((W, f), \Delta_X Y^Z) &\cong \C(W, Y^Z)\\
    &\cong \C(W \times Z, Y)\\
    &\cong (\C \downarrow X)((W \times (X \times Z), \langle f, p_1 \rangle), \Delta_X Y)\\
    &\cong (\C \downarrow X)((W, f) \times \Delta_X Z, \Delta_X Y)\\
  \end{align*}
\end{proof}

@@

Now, in $ \SET $, for a family $ (Y_x)_x $, consider the dependent product $ \prod_{x : X} Y_x $. Its elements $ (y_x)_x $ can be identified with morphisms from the terminal set: $ \{ \star \} \to \prod_{x : X} Y_x $, sending $ \star $ to $ (y_x)_x $. However, they can also be identified with the morphisms $ f: X \to \sum_{x : X} Y_x $ that make the following diagram commute, sending $ x $ to $ y_x $:
\begin{center}
  \begin{tikzcd}
    X \arrow[rd, "\id X"'] \arrow[rr, "f"] & & \sum_{x : X} Y_x \arrow[ld, "\pi_1"]\\
    & X
  \end{tikzcd}
\end{center}
These are morphisms in $ (\SET \downarrow X) $ from $ (X, \id X) $ to $ (\sum_{x : X} Y_x, \pi_1) $. Note that for the terminal morphism $ f: X \to \{ \star \} $, we have $ (X, \id X) = f^*(\{\star\}, \id{\{\star\}}) $. To summarize, we have a bijection
\[ (\SET \downarrow X)\left(f^*(\{ \star \}, \id{\{\star\}}), (Y_x)_x \right) \cong (\SET \downarrow {\{\star\}})\left( (\{\star\}, \id{\{ \star \}}), \left( \prod_{x : X} Y_x, ! \right) \right) \]
for $ ! $ the terminal morphism. Now, for an internal indexed family $ f : Y \to X $ and given a family of families $ ((Z_y)_{y : Y_x})_{x : X} $, we can wonder whether we can construct the family of dependent products $ (\prod_{y : Y_x} Z_y)_x $. In $ \SET $, this is definitely possible, and from this, we get a bijection again
\[ \left(\SET \downarrow \sum_{x : X} Y_x \right)\left(f^*(X, \id X), (Z_y)_{x, y} \right) \cong (\SET \downarrow X)\left((X, \id X), \left(\prod_{y : Y_x} Z_y\right)_x\right). \]
with $ p : \sum_{x : X} \sum_{y : Y_x} Z_y \to \sum_{x : X} Y_x $ defined as $ p(x, y, z) = (x, y) $. These bijections suggest an adjunction $ f^* \dashv \prod_{\dots} $. We can use this to define in general
\begin{definition}
  For a category $ \C $ and a morphism $ f: Y \to X $, the \iindex{dependent product} along $ f $ is, if it exists, the right adjoint to the pullback functor:
  \begin{center}
    \begin{tikzcd}
      (\C \downarrow X) \arrow[r, shift left=2, "f^*"{name=A}] &
      (\C \downarrow Y) \arrow[l, shift left=2, "\prod_f"{name=B}]
      \ar[from=A, to=B, symbol=\dashv]
    \end{tikzcd}
  \end{center}
\end{definition}
\begin{remark}
  As argued above, we can recover the familiar dependent product $ \prod_{x : X} Y_x $ of a family $ (Y_x)_x $ as the dependent product $ \prod_f (\sum_x Y_x, \pi_1) $ along the terminal morphism $ f: X \to I $. Here we use the equivalence between $ (\C \downarrow I) $ and $ \C $.
\end{remark}

Now we turn our attention to dependent sums. In $ \SET $, let $ (Y_x)_x $ and $ (Y^\prime_x)_x $ be two families over $ X $ and let $ ((Z_y)_{y : Y_x})_{x : X} $ be a family of families. Let $ f: \sum_{x : X} Y_x \to X $ be the internal representation of $ (Y_x)_x $. A family of maps $ g_x : (\sum_{y : Y_x} Z_y)_x \to Y^\prime_x $ consists of maps $ Z_y \to Y^\prime_x $ for all $ y : Y_x $, so these are maps $ g_y : Z_y \to Y^\prime_{f(y)} $. This gives a bijection
\[ (\SET \downarrow X)\left(\left( \sum_{y : Y_x} Z_y \right)_x, (Y^\prime_x)_x \right) \cong \left(\SET \downarrow \left(\sum_{x : X} Y_x \right) \right)((Z_y)_y, f^*((Y^\prime_x)_x)). \]
This, again, suggests an adjunction which we will use as a definition.
\begin{definition}
  For a category $ \C $ and a morphism $ f: Y \to X $, the \iindex{dependent sum} along $ f $ is, if it exists, the left adjoint to the pullback functor:
  \begin{center}
    \begin{tikzcd}
      (\C \downarrow X) \arrow[r, shift right=2, "f^*"'{name=B}] &
      (\C \downarrow Y) \arrow[l, shift right=2, "\sum_f"'{name=A}]
      \ar[from=A, to=B, symbol=\dashv]
    \end{tikzcd}
  \end{center}
\end{definition}

However, note that the conversion from an external to an internal representation in $ \SET $ already contained a dependent sum, which is no coincidence. It turns out that in practice, we will never have a hard time obtaining dependent sums:
\begin{lemma}\label{lem:sum-postcomposition}
  Let $ f : \C(Y, X) $ be a morphism in a category. If the pullback functor $ f^*: (\C \downarrow X) \to (\C \downarrow Y) $ exists, it has a left adjoint given by postcomposition with $ f $.
\end{lemma}
\begin{proof}
  For morphisms $ g: Z \to X $, $ h: W \to Y $, the universal property of the pullback, with the following diagram
  \begin{center}
    \begin{tikzcd}
      W \arrow[r, dashed, "\varphi"'] \arrow[rr, "\psi", bend left] \arrow[dr, "h"'] & f^* Z \arrow[d, "f^* g"'] \arrow[r] \arrow[rd, phantom, "\lrcorner", very near start] & Z \arrow[d, "g"]\\
      & Y \arrow[r, "f"'] & X
    \end{tikzcd}
  \end{center}
  gives a bijection between morphisms $ \varphi: W \to f^* Z $ that commute with $ h $ and $ f^* g $, and morphisms $ \psi: W \to Z $ that commute with $ h $, $ g $ and $ f $. In other words:
  \[ (\C \downarrow X)(h \cdot f, g) \cong (\C \downarrow Y)(h, f^*(g)), \]
  which shows the adjunction.
\end{proof}

@@

  Now, note that if we have a coequalizer $ W $ of $ \id Z $ and $ a $, and an equalizer $ Y $ of $ \id X $ and $ b $ (in particular, if $ W $ and $ Y $ are retracts), the universal properties of these give a bijection
  \[ \D(W, Y) \cong \{ f: \D(Z, Y) \mid a \cdot f = f \} \cong \{ f: \D(Z, X) \mid a \cdot f = f = f \cdot b \}, \]
  which exactly describes the morphisms in $ \bar C $.
  \begin{center}
    \begin{tikzcd}
      Z \arrow[r, shift left, "\id Z"] \arrow[r, shift right, "a"'] & Z \arrow[r] \arrow[d] \arrow[rd] & W \arrow[d]\\
      X & X \arrow[l, shift left, "\id X"] \arrow[l, shift right, "b"'] & Y \arrow[l]
    \end{tikzcd}
  \end{center}
\end{proof}

@@

\begin{corollary}\label{cor:karoubi-presheaf}
  As the category $ \SET $ is cocomplete, we have an equivalence between the category of presheaves on $ \C $ and the category of presheaves on $ \overline \C $:
  \[ [\op \C, \SET] \simeq [\overline{\op \C}, \SET] \simeq [\op{\overline \C}, \SET]. \]
\end{corollary}

### 2024-11-25 300-univalent-foundations.tex
Improve titles, make some sentences active instead of passive
Sort out simeq and cong

It is important to note that Martin-Löf type theory has \textit{identity types}: given a type $ T $ and elements $ x, y: T $, we have a type $ \mathrm{Id}_T(x, y) $ (note that this is a dependent type), which we will usually denote with $ x = y $. An element $ p: x = y $ is a proof that $ x $ is `equal' to $ y $. This type comes with an interesting induction principle called \iindex{path induction}: we can show any statement about paths $ p: x = y $ for generic $ x $ and $ y $, if we can show it about the `trivial' path $ \mathrm{refl}: x = x $ for generic $ x $. For example, symmetry of the equality boils down to a function $ \prod_{x y : T} x = y \to y = x $. We construct this using path induction with the function that sends $ \mathrm{refl}: x = x $ to itself. For more information, see \autocite[][Section 1.12.1]{hottbook}.

In set-theoretic mathematics, there is the concept of a `bijection' $ S \simeq T $ of sets (or an isomorphism in the category $ \SET $), which is often treated as an equivalence. It consists of functions $ f: S \to T $ and $ g: T \to S $ with $ f \cdot g = \id S $ and $ g \cdot f = \id T $. In type theory, we have a similar concept, which is called `equivalence' (of types) $ S \simeq T $. Since bijections are not well-behaved for types that are not sets (see Section \ref{sec:homotopy-props-sets}), because in those cases, a function $ f: S \to T $ can have multiple distinct inverses. Therefore, we define $ S \simeq T := \sum_{f: S \to T} \mathrm{isequiv}(f) $, for some predicate $ \mathrm{isequiv}: (S \to T) \to \TYPE $ \autocite[][Equation 2.4.10]{hottbook}. However, intuitively we can still think of these as bijections.

Using the identity type, we can make our statement of the univalence principle more precise (and a bit stronger). For types, we can construct a function
\[ \mathrm{idtoequiv}: \prod_{S, T: \TYPE} (S = T) \to (S \simeq T). \]
We construct this function using path induction with the identity bijection $ \prod_S \id S: (S \simeq S) $. In fact, this is a specific case of the following: for a category $ C $, if we denote the type of isomorphisms between objects $ c $ and $ d $ with $ c \cong d $, we can construct a function
\[ \mathrm{idtoiso}: \prod_{c, d: C} (c = d) \to (c \cong d), \]
using path induction with the identity isomorphism $ \prod_{c: C} \id c: (c \cong c) $. We can then formulate the univalence principle for categories as

@@

  Now, two initial functors $ F : C \to D $ and $ F^\prime : C \to D^\prime $ factor uniquely through each other:
  \[ F^\prime = F \bullet G \quad \text{and} \quad F = F^\prime \bullet G^\prime \]
  for two functors $ G $ and $ G^\prime $ and since $ F $ and $ F^\prime $ also factor uniquely through themselves, we have
  \[ G \bullet G^\prime = \id D \quad \text{and} \quad G^\prime \bullet G = \id{D^\prime}. \]
  From this, we can construct an equivalence $ D \cong D^\prime $. Since $ D $ and $ D^\prime $ are univalent, this equivalence gives an equality of categories, and because $ F \bullet G = F^\prime $, we have an equality of pairs
  \[ (D, F) = (D^\prime, F^\prime). \]
\end{proof}

@@

The last structure in this section for which we want to consider the univalence axiom, is the 2-category $ \Cat $ of categories. In general, we cannot show that this satisfies the univalence principle. However, we will restrict our attention to the sub-2-category of univalent categories, which are the categories that we want to study. Then Theorem 6.8 in \autocite{univalent-categories} shows that for univalent categories $ C $ and $ D $, there is an equivalence between $ C = D $ and $ C \simeq D $, where $ C \simeq D $ denotes the type of (adjoint) equivalences of categories (see Definition \ref{def:equivalence-of-categories}).

@@

\subsection{Caution: `transport hell'}
Now, it seems that we can transport all properties and data along equalities. And of course, that is true, but some caution needs to be taken with this when working with a proof assistant.

>>>>

It is important to note that Martin-Löf type theory has \textit{identity types}: given a type $ T $ and elements $ x, y: T $, we have a type $ \mathrm{Id}_T(x, y) $ (note that this is a dependent type), which we will usually denote with $ x = y $. An element $ p: x = y $ is a proof that $ x $ is `equal' to $ y $. This type comes with an interesting induction principle called \iindex{path induction}: we can show any statement about paths $ p: x = y $ for generic $ x $ and $ y $, if we can show it about the `trivial' path $ \mathrm{refl}: x = x $ for generic $ x $. Often, a way to think about this, is that if we have a proof of equality between $ x $ and $ y $, we can replace $ y $ by $ x $ in any statement. For example, symmetry of the equality boils down to a function $ \prod_{x y : T} x = y \to y = x $. We construct this using path induction with the function that sends $ \mathrm{refl}: x = x $ to itself. For more information, see \autocite[][Section 1.12.1]{hottbook}.

In set-theoretic mathematics, there is the concept of a `bijection' $ S \cong T $ of sets (or an isomorphism in the category $ \SET $), which is often treated as an equivalence. It consists of functions $ f: S \to T $ and $ g: T \to S $ with $ f \cdot g = \id S $ and $ g \cdot f = \id T $. In type theory, we have a similar concept, which is called `equivalence' (of types) $ S \simeq T $. Since bijections are not well-behaved for types that are not sets (see Section \ref{sec:homotopy-props-sets}), because in those cases, a function $ f: S \to T $ can have multiple distinct inverses. Therefore, we define $ S \simeq T := \sum_{f: S \to T} \mathrm{isequiv}(f) $, for some predicate $ \mathrm{isequiv}: (S \to T) \to \TYPE $ \autocite[][Equation 2.4.10]{hottbook}. However, intuitively we can still think of these as bijections.

Using the identity type, we can make our statement of the univalence principle more precise (and a bit stronger). For types, we can construct a function
\[ \mathrm{idtoequiv}: \prod_{S, T: \TYPE} (S = T) \to (S \simeq T). \]
We construct this function using path induction with the identity equivalence $ \prod_S \id S: (S \simeq S) $. More generally, for a category $ C $, if we denote the type of isomorphisms between objects $ c $ and $ d $ with $ c \cong d $, we can construct a function
\[ \mathrm{idtoiso}: \prod_{X, Y: C} (X = Y) \to (X \cong Y), \]
using path induction with the identity isomorphism $ \prod_{X: C} \id X: (X \cong X) $. If $ X $ and $ Y $ are sets, then $ \mathrm{idtoequiv} $ and $ \mathrm{idtoiso} $ coincide. We can then formulate the univalence principle for categories as

@@

  Now, two initial functors $ F : C \to D $ and $ F^\prime : C \to D^\prime $ factor uniquely through each other:
  \[ F^\prime = F \bullet G \quad \text{and} \quad F = F^\prime \bullet G^\prime \]
  for two functors $ G $ and $ G^\prime $ and since $ F $ and $ F^\prime $ also factor uniquely through themselves, we have
  \[ G \bullet G^\prime = \id D \quad \text{and} \quad G^\prime \bullet G = \id{D^\prime}. \]
  From this, we can construct an equivalence $ D \simeq D^\prime $. Since $ D $ and $ D^\prime $ are univalent, this equivalence gives an equality of categories, and because $ F \bullet G = F^\prime $, we have an equality of pairs
  \[ (D, F) = (D^\prime, F^\prime). \]
\end{proof}

@@

The last structure in this section for which we want to consider the univalence principle, is the 2-category $ \Cat $ of categories. In general, we cannot show that this satisfies the univalence principle. However, we will restrict our attention to the sub-2-category of univalent categories, which are the `well-behaved' categories in some sense. Then Theorem 6.8 in \autocite{univalent-categories} shows that for univalent categories $ C $ and $ D $, there is an equivalence between $ C = D $ and $ C \simeq D $, where $ C \simeq D $ denotes the type of (adjoint) equivalences of categories (see Definition \ref{def:equivalence-of-categories}).

@@

\subsection{Caution: `Transport Hell'}
Now, it seems that we can transport all properties and data along equalities. And of course, that is true, but some caution needs to be taken with this when working with a proof assistant.
